---
title: "a3_task3_Teague_Tran"
author: "Teague Tran"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(kableExtra)
library(lubridate)
#Read in the .csv file
urchins <- read_csv(here("data", "urchins.csv"))
```

##Section 1: Size comparisions: purple vs red urchins

#### ii Exploratory data visualization:
```{r, echo = TRUE, message = FALSE, warning = FALSE}
#create faceted histogram to see spread of purple vs red urchin size

ggplot(data = urchins, aes(x = SIZE)) +
  geom_histogram()+
  facet_wrap("COMMON_NAME")

#create box plot with common name on x-axis and size on y-axis

ggplot(data = urchins, aes(x = COMMON_NAME, y = SIZE))+ 
  geom_boxplot()+
  labs(x = "Species", y = "Size (cm)")
  
```

#### ii) Descriptive statistics table:
```{r, echo = TRUE, message = FALSE, warning = FALSE}
#Put summary statistics for sea urchin size into a summary table

summary_urchins <- urchins %>% 
  group_by(COMMON_NAME) %>% 
  summarize(mean_size = mean(SIZE),
            median_size = median(SIZE),
            sd_size = sd(SIZE),
            n = n())
# make table look a bit better, must find out how to separate better
kable(summary_urchins, booktabs = T, caption = "Urchin size summary statistics",
      col.names = c("Common name ",
                " mean size (cm) ", 
                "  median size (cm)  ",
                "  standard deviation (cm)  ",
                "urchin count  "))
  
```
#### iii) Takeaways

Both purple and red urchin have normally distributed data along a point, with red urchins having a wider spread (more variation). However, there is a 51.6% difference (~2.31 cm difference) in the mean and a 55% difference in median diameter (2.5cm difference). 

###### The percent difference was found by subtracting the purple urchin statistic from the red urchin statistic and dividing it by the purple urchin statistic.

## Section 2: Explore purple urchin counts by site and year

#### i)
```{r, echo = TRUE, message = FALSE, warning = FALSE}

purple_urchins <- urchins %>%
  #filter to only purple urchins
  filter(COMMON_NAME == "Purple Urchin") %>% 
  #convert date to class date
  mutate(date_new = mdy(DATE)) %>% 
  #create new  "year" column with just years
  mutate(year = year(date_new)) %>% 
  #Create new column with full names of sites
  mutate(site_full = (case_when(
  SITE == "NAPL" ~ "Naples Reef",
  SITE == "AQUE" ~ "Arroyo Quemado Reef",
  SITE == "CARP" ~ "Carpinteria Reef",
  SITE == "MOHK" ~ "Mohawk Reef",
  SITE == "IVEE" ~ "Isla Vista Reef"
    
  )
  )
  )
```
#### ii)



```{r, echo = TRUE, message = FALSE, warning = FALSE}
#Make a summary table containing the total number of purple urchins observed by site and year

urchin_table <- purple_urchins %>% 
  count(site_full, year)

urchin_table  
  
```

#### iii)
```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(urchin_table, aes(x = year, y = n)) +
  geom_col()+ 
  facet_wrap("site_full")+
  labs(x = "Year", y = "Count (# purple urchin)", title = "Count of Purple Urchins per year at each site")
```






