---
title: "a3_task4_Teague_Tran"
author: "Teague Tran"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(kableExtra)
library(lubridate)
#Read in the .csv file
urchins <- read_csv(here("data", "urchins.csv"))
```
#### Compare mean size of purple urchins in CONTROL and ANNUAL sections of Naples Reef in January 2010

```{r,echo = TRUE, message = FALSE, warning = FALSE}
#Code taken from task 3 that selects purple urchins + creates year + full site column that will be used in this section  ALSO ADDED MONTH THAT WAS NOT IN TASK 3.

purple_urchins <- urchins %>%
  #select only purple urchins
  filter(COMMON_NAME == "Purple Urchin") %>% 
  #convert date to class date
  mutate(date_new = mdy(DATE)) %>% 
  #create new  "year" column with just years
  mutate(year = year(date_new)) %>% 
  #create new "month" column with just months (WAS NOT IN TASK 3)
  mutate(month = month(date_new)) %>% 
  #Create new column with full names of sites
  mutate(site_full = (case_when(
  SITE == "NAPL" ~ "Naples Reef",
  SITE == "AQUE" ~ "Arroyo Quemado Reef",
  SITE == "CARP" ~ "Carpinteria Reef",
  SITE == "MOHK" ~ "Mohawk Reef",
  SITE == "IVEE" ~ "Isla Vista Reef"
    
  )
  )
  )
```
```{r}
#select Naples Reefs observations of purple urchins in January 2010 + treatment
jan2010_napurch <- purple_urchins %>%
  #filter out year = 2010, month = January (1), sites = Naples Reef
  filter(year == "2010", month == "1", site_full == "Naples Reef" )  %>% 
  #select columns "SIZE", "TREATMENT", "year", and "site_full"
  select("SIZE", "TREATMENT", "year","month","site_full")
  
```
#### Looking at histograms and QQ plots for each treatment to assess normality
```{r}
#creating a histogram of size and treatment to assess normaility
ggplot(data = jan2010_napurch, aes(x = SIZE)) +
  geom_histogram(bins = 15)+ 
  facet_wrap("TREATMENT")
```
```{r,echo = TRUE, message = FALSE, warning = FALSE}
#creating a QQ-plot to assess normaility
ggplot(data = jan2010_napurch, aes(sample = SIZE)) +
      geom_qq()+
      facet_wrap(~TREATMENT)
```


The histogram shows that there are higher counts in the middle, while the QQ plots show a linear relationship in both treatments. This means that both histogram and QQ plot show that there is normal distribution in both annual and control treatments.


#### Create a table of basic descriptive statistics for treatments
```{r,echo = TRUE, message = FALSE, warning = FALSE }
#summary statistics including mean, median, standard deviation, and sample size
urchin_summary <- jan2010_napurch %>% 
  group_by(TREATMENT) %>% 
  summarize(mean = mean(SIZE),
            median = median(SIZE),
            sd = sd(SIZE),
            n = n())
#show table
urchin_summary
```

#### Perform a t-test looking at the means between the two treatments

```{r,echo = TRUE, message = FALSE, warning = FALSE}
# perform t-test on mean treatments
seaurchin_ttest <- jan2010_napurch %>% 
  group_by(TREATMENT)
#separate out control
seaurchin_control <- seaurchin_ttest %>% 
  filter(TREATMENT == "CONTROL") %>% 
  pull(SIZE)
#separate out annual
seaurchin_annual <- seaurchin_ttest %>% 
  filter(TREATMENT == "ANNUAL") %>% 
  pull(SIZE)
#perform t-test on mean sizes
t.test(x= seaurchin_control, y = seaurchin_annual)
 
```

The t-test shows that the means from the samples are close to each other (difference = .190861), which means that it is more likely that the null hypothesis (difference in means = 0) is true. The null hypothesis seems more likely to be true when you take into account that the difference in means is only approximately 5%. A p-value of .25 means that there is a 25% chance that you will get means that have a difference of at least .190861 if the null hypothesis is true.


% difference was calculated by subtracting mean control by mean annual and dividing by mean annual.














